Explore CNV distributions of reference cells: SCPCP000015

### 2025-05-28 analyses/infercnv-consensus-cell-type/exploratory-notebooks/02_ewings-reference-cnv.Rmd
add endo-immune

This function prepares a TSV of `inferCNV` results by creating a `total_cnv_per_cell` column to summarize CNVs per cell.

@@

@@

```{r}
# inferCNV output files
infercnv_files <- list.files(
  path = result_dir,
  pattern = "_cnv-metadata\\.tsv",
  recursive = TRUE,
  full.names = TRUE
) |>
  # remove any containing "endo_immune" or "internal"; we only want pooled endo or immune
  purrr::discard(\(x) {
    stringr::str_detect(x, "endo-immune|internal")
  }) |>
  # name by "library-reference"
  purrr::set_names(\(x) {
    library_id <- stringr::str_split_i(basename(x), "_", 1)
  })
#    ref_name <- stringr::str_remove(basename(dirname(x)), "_pooled")
#    glue::glue("{library_id}-{ref_name}")
# })

@@

```{r, fig.height = 8}
immune_df <- infercnv_df |>
  dplyr::filter(normal_reference == "immune_pooled")

endo_df <- infercnv_df |>
  dplyr::filter(normal_reference == "endo_pooled")

ggplot(immune_df) +
  aes(x = total_cnv_per_cell, y = label, fill = label) +
  ggridges::geom_density_ridges2(scale = 0.9) +
  facet_wrap(vars(infercnv_library_id)) +
  theme(legend.position = "none") +
  ggtitle("Immune pooled reference")

ggplot(endo_df) +
  aes(x = total_cnv_per_cell, y = label, fill = label) +
  ggridges::geom_density_ridges2(scale = 0.9) +
  facet_wrap(vars(infercnv_library_id)) +
  theme(legend.position = "none") +
  ggtitle("Endo pooled reference")
```

* The `endo` reference cell CNV distributions generally do not overlap with tumor cell CNV distributions.
* The `immune` reference cell CNV distributions do have a wider spread but do not _substantially_ overlap with the tumor cell CNV distributions.

@@

```{r}
ref_immune_df <- immune_df |>
  dplyr::filter(cell_group == "reference") |>
  tidyr::separate(cell_id, into = c("reference_library_id", "barcode"), sep = "-", remove = FALSE)

>>>>

This function prepares a table of `inferCNV` results by creating a `total_cnv_per_cell` column to summarize CNVs per cell.

@@

This function plots ridge plots of total CNV values across cell groupings for a given normal reference.

```{r}
plot_cnv_distributions <- function(df, reference_name) {
  filtered_df <- infercnv_df |>
    dplyr::filter(normal_reference == reference_name)

  ggplot(filtered_df) +
    aes(x = total_cnv_per_cell, y = label, fill = label) +
    ggridges::geom_density_ridges2(scale = 0.9) +
    facet_wrap(vars(infercnv_library_id)) +
    theme(legend.position = "none") +
    ggtitle(reference_name)
}
```

@@

```{r}
# inferCNV output files
infercnv_files <- list.files(
  path = result_dir,
  pattern = "_cnv-metadata\\.tsv",
  recursive = TRUE,
  full.names = TRUE
) |>
  # remove any containing "internal"; we only want pooled
  purrr::discard(\(x) {
    stringr::str_detect(x, "internal")
  }) |>
  # name by "library-reference"
  purrr::set_names(\(x) {
    library_id <- stringr::str_split_i(basename(x), "_", 1)
  })

@@

```{r, fig.height = 8}
plot_cnv_distributions(infercnv_df, "endo_pooled")
plot_cnv_distributions(infercnv_df, "immune_pooled")
plot_cnv_distributions(infercnv_df, "endo-immune_pooled")
```

* The `endo` reference cell CNV distributions generally do not overlap with tumor cell CNV distributions.
* The `endo-immune` and `immune` reference cell CNV distributions do have a wider spread but do not _substantially_ overlap with the tumor cell CNV distributions.

@@

```{r}
ref_immune_df <- infercnv_df |>
  dplyr::filter(
    normal_reference == "immune_pooled",
    cell_group == "reference"
  ) |>
  tidyr::separate(cell_id, into = c("reference_library_id", "barcode"), sep = "-", remove = FALSE)
